### 2022-03-02 docs/report.Rmd
text tweaks

---
title: "Real-time epidemiological estimates from ONS Community Infection Survey data"
author: Sam Abbott, Sebastian Funk
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output: html_document
---

```{r setup, echo = FALSE, include = FALSE}
library(cowplot)
library(here)
library(magick)

estimates <- c(national = "National",
               regional = "Regional",
               age_school = "Age")

quantities <- c(est_prev = "PCR+ prevalence",
                infections = "Infection incidence",
                r = "Growth rate",
                Rt = "R",
                cumulative_infections = "Cumulative incidence",
                cumulative_exposure = "Proportion ever infected")

time_horizon <- c(est_prev = "1year",
                  infections = "1year",
                  r = "3months",
                  Rt = "3months",
                  cumulative_infections = "1year",
                  cumulative_exposure = "1year")

additional_text <- list(
  est_prev = "Proportion testing PCR positive.",
  infections = "Daily proportion newly infected.",
  r = "Growth rate of new infections.",
  Rt = "Infection reproduction number.",
  cumulative_infections = "Cumulative incidence of infections, including reinfections.",
  cumulative_exposure = "Crude estimate of the number of people ever infected under the extremely simplifying assumption that the risk of infection is independent of prior infection or vaccination status."
)
```

Estimates were derived from the modelled daily and weekly data published weekly as part of the Covid-19 Community Infection Survey by the Office of National Statistics. Prevalence and incidence estimates are shown for the last year, and reproduction number for the past three months. Code to reproduce the results and links to data tables with estimates are available at [https://github.com/epiforecasts/inc2prev](https://github.com/epiforecasts/inc2prev).

```{r, results = 'asis', echo = FALSE}
for (level in names(estimates)) {
  cat("\n\n#", estimates[[level]], "estimates", "{.tabset}", "\n")
  for (quantity in names(quantities)) {
    cat("\n\n##", quantities[[quantity]], "\n")
    cat(additional_text[[quantity]], "\n\n")
    file_name <-
      paste0(level, "_", quantity, "_", time_horizon[[quantity]], ".png")
    p <- ggdraw() + draw_image(here::here("figures", "additional", file_name))
    print(p)
  }
}
```

# License

This project uses data from the Office for National Statistics Community Infection Survey, which is licensed under the [Open Government License v3.0](https://www.ons.gov.uk/peoplepopulationandcommunity/healthandsocialcare/conditionsanddiseases/datasets/coronaviruscovid19infectionsurveydata).